---
title: "Breast Cancer Machine Learning Report"
author: "Samuel O'Ferrell"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(ggplot2)
```

### Reading in Data

```{r}
data <- read_csv("breastcancer.csv")
```

### Data Discussion

**Additional Information**

Features are computed from a digitized image of a fine needle aspirate (FNA) of a breast mass. They describe characteristics of the cell nuclei present in the image. A few of the images can be found at http://www.cs.wisc.edu/\~street/images/ Separating plane described above was obtained using Multisurface Method-Tree (MSM-T) \[K. P. Bennett, "Decision Tree Construction Via Linear Programming." Proceedings of the 4th Midwest Artificial Intelligence and Cognitive Science Society, pp. 97-101, 1992\], a classification method which uses linear programming to construct a decision tree. Relevant features were selected using an exhaustive search in the space of 1-4 features and 1-3 separating planes. The actual linear program used to obtain the separating plane in the 3-dimensional space is that described in: \[K. P. Bennett and O. L. Mangasarian: "Robust Linear Programming Discrimination of Two Linearly Inseparable Sets", Optimization Methods and Software 1, 1992, 23-34\].

<https://archive.ics.uci.edu/dataset/17/breast+cancer+wisconsin+diagnostic>

#### Variables

-   id

-   diagnosis - M for Malignant or B for Benign

-   radius_mean - Mean Radius of Lobes

-   texture_mean - Mean Standard Deviation of Gray-Scale Values of Tumor

-   perimeter_mean - Mean Outer Perimeter of Tumor

-   area_mean - Mean Area of Lobes

-   smoothness_mean - Mean Smoothness of the Tumor

-   compactness_mean - Mean Compactness of the Tumor

-   concavity_mean - Mean of Concavity of the Tumor

-   concave points_mean - Mean Number of Concave Points on the Tumor

-   symmetry_mean - Mean Symmetry of the Tumor

-   fractal_dimension_mean - Mean Fractal Dimension of Tumor

-   radius_se - SE of Radius

-   texture_se - SE of Texture

-   perimeter_se - SE of Perimeter

-   area_se - SE of Area

-   smoothness_se - SE of Smoothness

-   compactness_se - SE of Compactness

-   concavity_se - SE of Concavity

-   concave points_se - SE of Concave Points

-   symmetry_se - SE of Symmetry

-   fractal_dimension_se - SE of Fractal Dimension

-   radius_worst - Worst Radius

-   texture_worst - Worst Texture

-   perimeter_worst - Worst Perimeter

-   area_worst - Worst Area

-   smoothness_worst - Worst Smoothness

-   compactness_worst - Worst Compactness

-   concavity_worst - Worst Concavity

-   concave points_worst - Worst Concave Point

-   symmetry_worst - Worst Symmetry

-   fractal_dimension_worst - Worst Fractal Dimension

    1\) ID number
    2) Diagnosis (M = malignant, B = benign)
    3-32)

    Ten real-valued features are computed for each cell nucleus:

    	a) radius (mean of distances from center to points on the perimeter)
    	b) texture (standard deviation of gray-scale values)
    	c) perimeter
    	d) area
    	e) smoothness (local variation in radius lengths)
    	f) compactness (perimeter\^2 / area - 1.0)
    	g) concavity (severity of concave portions of the contour)
    	h) concave points (number of concave portions of the contour)
    	i) symmetry
    	j) fractal dimension ("coastline approximation" - 1)

### Understanding Data

Let's look at the summary of the data:

```{r}
summary(data)
```

Checking for missing data:

```{r}
colSums(is.na(data))
```

Here we can see we have no missing data.

### Data Exploration

Now, lets make some plots to visualize the data:

```{r}
table(data$diagnosis)
```
